# 

# Dataset for diamond

sourse: from tidyverse library

```{r}
# import the necessary library
library(tidyverse)
library(ggplot2)

```

# Loading the dataset

```{r load_data, message=TRUE}
diamonds_df = read_csv('./datasets/diamonds.csv') # diamonds datasets 
print(diamonds_df)
```

# Exploring the dataset.

while exploring the dataset, we explore and we start by knowing the following:

-   Data Type of each column

-   Dimension of the dataset (Number of rows and column)

-   Number/percentage of missing values.

-   Descriptive summary(statistics) of the dataset in general.

# Data Type of each column

```{r}

str(diamonds)
```

```{r}
# dimension of the datasets
rc <- dim(diamonds_df)
cat('Row -', rc[1], '\n')
cat('column -', rc[2])
```

```{r}
#datatype of each variables
sapply(diamonds_df, class)
```

# Number/percentage of missing values

```{r}
colSums(is.na(diamonds_df))

```

#### Descriptive summary (statistics) of the datasets in general

```{r}
options(digits = 2)
desc_sum <- summary(diamonds_df)
print(desc_sum)
```

### Transforming the dataset

```{r}
# basic data cleaning and conversion to integers
diamonds <- diamonds_df %>% 
   mutate(across(c(carat, depth,table, x, y, z), as.numeric)) %>% 
      filter_all(all_vars(!is.na(.))) # remove rows with any missing value
```

```{r}
head(diamonds_df)
```

# Exploratory analysis

```{r}
# counting the numbers of the cut category
count_cut = diamonds_df %>% 
   count(cut)
print(count_cut)
```

```{r}
# counting the numbers of the colour category
count_color = diamonds_df %>% 
   count(color)
print(count_color)
```

```{r}
# counting the numbers of the clearity category
count_clarity = diamonds_df %>% 
   count(clarity)
print(count_clarity)
```

#### VISUALIZATION

```{r}
cut_data <- ggplot(diamonds_df, aes(x = cut, fill = cut))  +
   geom_bar() +
   labs(title = "Distribution of cut categories") +
   theme_minimal()

cut_data

ggsave('./image_plot/cut_data.png', cut_data + theme(panel.background = element_rect(fill = 'white')), width = 12, height=7)
```

```{r}
clarity_data <- ggplot(diamonds_df, aes(x = clarity, fill = clarity))  +
   
   geom_bar() +
   labs(title = "Distribution of clarity categories") +
   theme_dark()

clarity_data
ggsave('./image_plot/cut_data.png', clarity_data + theme(panel.background = element_rect(fill = 'white')), width = 12, height=7)
```

```{r}
color_data <- ggplot(diamonds_df, aes(x = color, fill = color))  +
   geom_bar() +
   labs(title = "Distribution of color categories") +
   theme_minimal()
color_data

ggsave('./image_plot/cut_data.png', cut_data + theme(panel.background = element_rect(fill = 'white')), width = 12, height=7)
```

Data Questions to analyze

-   What is the distribution of carat weights in the dataset, and how does it correlate with the price of diamonds?

-   How are the prices distributed among different cut categories?

-   Is there a noticeable trend between the depth percentage and the price of diamonds?

-   Which combination of cut, color, and clarity tends to have the highest average price?

-   Are there any outliers in the dataset, and how do they impact the overall distribution of prices?

#### solution

```{r}
# scatter plot  of carat vs. price
ggplot (diamonds_df, aes(x = carat, y = price, fill = carat)) +
   geom_point(alpha  = 0.5) +
   labs(title = "scatter plot of carat vs. price", 
        x = "carat",
        y = "price") +
      theme_minimal()
```

### interpretation

## This is to say there is a correlation between the carat and the price, i.e the increase in carat is a result to the increase in price.

```{r}
# boxplot of prices by cut
ggplot (diamonds_df, aes(x= cut, y = price, fill = cut)) +
   geom_boxplot() +
   labs(title = "boxplot of prices by cut", 
        x = "cut",
        y = "price") +
   theme_minimal()
```

## Interpretation

#### from the graph above we can say the ideal cut is

```{r}
#
ggplot(diamonds_df, aes(x = depth, y = price,)) +
   geom_point(alpha  = 0.5) +
   geom_smooth(method = "lm", se = FALSE, color = "blue") +
   labs(title = "scatter plot of depth vs. price regression line", 
        x = "depth",
        y = "price") +
      theme_minimal()
```

## Interpretation

from the graph above this says there is no correlation at all between price and depth

```{r}
zero <- diamonds_df %>% 
   filter(price  > 2000)

zero
```

```{r}
ggplot(diamonds_df, aes(x = table, y = price)) + 
   geom_point(alpha = 0.5) +
   labs(title = "relationship betwwen the table and price of the diamonds",
        x = "table",
        y = "price") +
   theme_minimal()
```

```{r}
average_prices <- diamonds_df %>%
   group_by(cut, color, clarity) %>%
   summarize(avg_price = mean(price))

top_combinations <- head(arrange(average_prices, desc(avg_price)), 5)
print(top_combinations)
```

```{r}
ggplot(diamonds_df, aes(y = price)) + 
   geom_boxplot() +
   labs(title = "boxplot of prices",
        y = "price") +
   theme_minimal()
```
