# Machine Learning Model Building

#### Agenda

1.  Loading Necessary Liraries
2.  Load the datasets
3.  selectin of relevant variables for abstraction and Generalization.
4.  Feature Engeneering (for accurate modeling).
5.  Datasets spliting (traininng and testing)
6.  Modeling Building(with Linear Regression, Random forest and Decision Tree algorithms)

```{r}
# load necessary libraries
library(tidyverse)
library(caret)
library(randomForest)
library(corrplot)

```

```{r}
diamonds_df <- read_csv ("dataset/diamonds.csv")
```

```{r}

diamonds_df[, c("carat", "cut",  "color", "clarity", "depth", "table", "price")]

```

## Correlation Matrix

```{r}
correlation_matrix <- cor(
   diamonds_df[, c("carat",  "depth", "x", "y", "z","table", "price" )]
)
```

```{r}
print("Correlation Matrix:")
print(correlation_matrix)

```

```{r}
corrplot (correlation_matrix, method = "color", addCoef.col = "black")
```

```{r}
diamonds_selected <- diamonds_df [, c("carat", "cut", "color", "clarity", "x", "y", "z", "price")]

diamonds_selected
```

# splitting the datasets into training and testing

```{r}
set.seed(62)
train_index <- createDataPartition(diamonds_selected$price, p = 0.7, list = FALSE,)
train_data <- diamonds_selected[train_index, ]
test_data <- diamonds_selected[- train_index, ]
```

```{r}
test_data
```

## Model Building  Using the linear Regression Algorithm

Linear regression is a statistical method used to model the relationship between a dependant variable (target) and one or more independent variables (features) by fitting a linear equation to observed data. it asumes that the relationship between the variables is approximately linear.

```{r}
# Linear Regression
linear_model <- lm(price ~ ., data = train_data)
```

```{r}


# Evaluate Linear Regression model
linear_predictions_train <- predict(linear_model, newdata = train_data)
linear_mse_train <- mean((train_data$price - linear_predictions_train)^2)
cat("\nLinear Regression - Mean Squared Error (Train):", linear_mse_train, "\n")

linear_predictions_test <- predict(linear_model, newdata = test_data)
linear_mse_test <- mean((test_data$price - linear_predictions_test)^2)
cat("Linear Regression - Mean Squared Error (Test):", linear_mse_test, "\n")


```
